---
title: "Exploring multiple dimensions of biodiversity in Japanese ferns and lycophytes"
bibliography: "references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
    includes:
      in_header: newphyt.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  chunk_output_type: console
# nocite: | 
#   @Vogel1968, @Wollenweber2000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
```

```{r load-packages, cache = FALSE}
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("code/functions.R"))

# Load targets
loadd(list = c(
  "all_cells",
  "occ_data_ferns",
  "occ_data_ferns_north",
  "occ_data_pteridos",
  "alpha_div_pteridos",
  "world_map"), 
  cache = ja_fern_cache)
```

<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
\pagenumbering{gobble}

<!-- Set left justification (better conversion to Word than default full justification) -->
\raggedright

<!-- Don't indent authors -->
\setlength{\parindent}{0ex}

Joel H. Nitta^1^ and Atsushi Ebihara^2^

^1^Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.

^2^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan  

Author for correspondence:  
*Joel H. Nitta*  
*Tel: +1 6174960515*  
*Email: joelnitta@gmail.com*

<!-- Set indent to about 3 characters hereafter -->
\setlength{\parindent}{3ex}

\clearpage

## Abstract

The pteridophytes (ferns and lycophytes) of Japan are one of the best-characterized regional floras in the world thanks to decades of collection and taxonomic work. Here, we combine phylogenetic, functional, and environmental datasets to characterize the spatial distribution of biodiversity in the pteridophytes of Japan.

**Key words:** biodiversity, biogeography, phylogenetic diversity. 

\clearpage

## Introduction

## Materials and Methods

### Study Site

Our data include all ferns and lycophytes known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E). Japan consists of four main islands and thousands of smaller ones. Our data exlude islands smaller than XX km^2^. The climate is subtropical in the south and highly seasonal in other parts of the country.

### Occurrence data

We used occurrence data of ferns and lycophytes in Japan compiled from > 200,000 vouchered specimens including 721 non-hybrid taxa (species, subspecies, and varieties) and 371 interspecific hybrids (1090 taxa total) [@Ebihara2019b]. The data are formatted as taxon presences in 10 km × 10 km grid cells (hereafter '10 km cells'). The 10 km grid cells correspond to the secondary level of basic grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) (`r nrow(all_cells) %>% comma` 10 km cells total).

### Trait data

We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b]. Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens. All qualitative traits formatted in binary format; e.g., a single trait with three states is formatted as three binary traits. Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately. Quantitative traits include aspects of leaf size (e.g., frond length, frond width, stipe length, etc.). Qualitative traits include aspects of leaf morphology (e.g., margin shape, leaf color, indusia shape, spore arrangement, etc.).

### Statistical analysis

We used two distance-based metrics to quantify alpha biodiversity, mean phylogenetic distance (MPD; the mean distance between all co-occurring taxa in a community) and mean nearest taxon distance (MNTD; the mean distance between closest taxon pairs in a community) [@Webb2000; @Webb2002]. Although MPD and MNTD were originally formulated for phylogenetic distances, they can be applied to any type of distance matrix. We calculated phylogenetic distances from the phylogenetic tree using the "cophenetic" function in R. We calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011]. We weighted quantitative and qualitative traits equally, and weighted each binary component trait of each qualitative trait equally.

Since observed values of MPD and MNTD are known to correlate with species richness [@Swenson2014], we determined whether communities are overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD. We compared observed values to those of a null distribution of 999 random communities with the \function{ses.mpd} and \function{ses.mntd} functions in the \package{picante} R package [@Kembel2010]. Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities. Positive values of SES of MPD and MNTD indicate overdispersion, and negative values indicate clustering. We hereafter refer to the SES of phylogenetic and functional MPD and MNTD as MPD~phy~ and MNTD~phy~, and MPD~func~ and MTND~func~, respectively.

A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude. When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering. Lycophytes, which occur on long branches, may have a similar effect. Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_NS").

### Data accessibility

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

### Richness

Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (\fig \@ref(fig:div-pteridos)a). Richness tended to increase with elevation ((\fig \@ref(fig:div-el)a).

### Phylogenetic diversity

Standard effect sizes of mean phylogenetic diversity (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) were either not correlated or only very weakly correlated with species richness (SI Fig. ), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness.

Values of MPD~phy~ are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan (\fig \@ref(fig:div-pteridos)b), with similar patterns between pteridophytes and ferns (SI Fig. ). The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispersion on the islands of Amami-oshima in sourthern Japan reverses to clustering (SI Fig. ).

The signal of clustering is much stronger when measured with MNTD~phy~, especially in mountainous areas of central Honshu (\fig \@ref(fig:div-pteridos)c). Patterns are similar between all three datasets except that MNTD~phy~ values are smaller in the "ferns_NS" dataset (SI Fig. ).

Values of MPD and MNTD decrease significantly with elevation across all datasets (\fig \@ref(fig:div-el)b, c; linear model).

### Functional diversity

Spatial patterns of functional diversity differed from phylogenetic diversity and by metric. When measured with MPD~func~, pteridophytes were overdispersed in Hokkaido, northern Honshu, and the smaller islands south of Kyushu, but clustered in southern Honshu, Shikoku, and Kyushu (\fig \@ref(fig:div-pteridos)d). Less spatial structure was evident in MNTD~func~, with more clustering overall (\fig \@ref(fig:div-pteridos)e).

Functional diversity increased with elevation when measured with MPD~func~, but not MNTD~func~ (\fig \@ref(fig:div-el)d, e; linear model)

### Breeding system

Sexual diploid taxa tend to be distributed in northern latitudes (\fig \@ref(fig:div-pteridos)f).

Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (linear model). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

## Discussion

## Acknowledgements

We thank people

## Author Contribution

Who did what

## References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

\clearpage

## Supporting Information

Additional Supporting Information may be found online in the Supporting Information section at the end of the article:

**Table S1** 

**Fig. S1**

\clearpage

**\fig \@ref(fig:div-pteridos).** Alpha diversity of ferns and lycophytes in Japan. **a** species richness. **b** Phylogenetic mean phylogenetic distance (MPD~phy~). **c** Phylogenetic mean nearest taxon distance (MNTD~phy~). **d** Morphological MPD (MPD~morph~). **e** Morphological MNTD (MNTD~morph~). 

**\fig \@ref(fig:div-el).** Relationship between alpha diversity metrics and elevation (m) in native Japanese pteridophytes. Blue trendline shows significant linear model fit with confidence interval in pink. **a** species richness. **b** Phylogenetic mean phylogenetic distance (MPD~phy~). **c** Phylogenetic mean nearest taxon distance (MNTD~phy~). **d** Morphological MPD (MPD~morph~). **e** Morphological MNTD (MNTD~morph~). 

\clearpage

<!-- Tables -->

\clearpage

<!-- Figures -->

```{r div-pteridos, fig.height = 9, fig.width = 7, fig.cap = "Alpha diversity of ferns and lycophytes in Japan"}

# pterido richness
  a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "richness", 
    metric_title = "No. species ",
    label = "a") +
    scale_fill_scico(palette = "bamako", na.value="grey")
  
  # pterido_phy_mpd
  b <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mpd_obs_z", 
    metric_title = "SES of MPD",
    label = "b"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))
  
  # pterido_phy_mntd
  c <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mntd_obs_z", 
    metric_title = "SES of MNTD",
    label = "c"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))
  
  # pterido_func_mpd
  d <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "func_mpd_obs_z", 
    metric_title = "SES of MPD",
    label = "d"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))
  
  # pterido_func_mntd
  e <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "func_mntd_obs_z", 
    metric_title = "SES of MNTD",
    label = "e"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))
  
  f <- make_diversity_map(
  div_data = alpha_div_pteridos, 
  world_map = world_map, 
  occ_data = occ_data_pteridos, 
  div_metric = "percent_sex_dip",
  metric_title = "% sex. diploid",
  label = "f"
)  +
    scale_fill_scico(palette = "bamako", na.value="grey")
  
  a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)

ggsave(here::here("ms/fig_1.pdf"), width = 7, height = 9, units = "in")
```

```{r div-el, fig.height = 3.5, fig.width = 7, fig.cap = "Relationship between alpha diversity metrics and elevation"}
a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "richness", "Elevation (m)", "Richness") +
  labs(title = "a", x = "Elevation (m)")
b <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPDphy") +
  labs(title = "b", x = "Elevation (m)")
c <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTDphy") +
  labs(title = "c", x = "Elevation (m)")
d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPDfunc") +
  labs(title = "d", x = "Elevation (m)")
e <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTDfunc") +
  labs(title = "e", x = "Elevation (m)")
a + b + c + d + e + plot_layout(ncol = 3, nrow = 2)

ggsave(here::here("ms/fig_2.pdf"), width = 7, height = 3.5, units = "in")
```